% PROJECT: <ETD> Electronic Thesis and Dissertation Initiative
%   TITLE: LaTeX report template for ETDs in LaTeX
%  AUTHOR: Neill Kipp, nkipp@vt.edu
%     URL: http://etd.vt.edu/latex/
% SAVE AS: etd.tex
% REVISED: September 6, 1997 [GMc 8/30/10]

% Instructions: Remove the data from this document and replace it with your own,
% keeping the style and formatting information intact.  More instructions
% appear on the Web site listed above.

%\documentclass[12pt,dvips]{report}
\documentclass[12pt]{report}
\setlength{\textwidth}{6.5in}
\setlength{\textheight}{8.5in}
\setlength{\evensidemargin}{0in}
\setlength{\oddsidemargin}{0in}
\setlength{\topmargin}{0in}
\setlength{\parindent}{0pt}
\setlength{\parskip}{0.1in}
% Uncomment for double-spaced document.
% \renewcommand{\baselinestretch}{2}
% \usepackage{epsf}

%--------packages I used----------------------------
\usepackage{verbatim}
\usepackage{latexsym}
\usepackage{amsmath, amsfonts, amssymb}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subfig}
\usepackage{url}
\usepackage{color}
\newcommand{\aravind}[1]{[{\color{blue}\textit{Aravind}\textit{: #1}}]}
\newcommand{\naren}[1]{[{\color{red}\textit{Naren}\textit{: #1}}]}
\newcommand{\TODO}[2]{{\color{magenta}[\textbf{TODO:} #2 \\ \textbf{Assignee:} #1]}}


% symbols and stuff from Bert's paper
\newcommand{\then}{\Rightarrow}
\newcommand{\softor}{\operatornamewithlimits{\tilde{\vee}}}
\newcommand{\softand}{\operatornamewithlimits{\tilde{\wedge}}}
\newcommand{\softthen}{\operatornamewithlimits{\tilde{\then}}}
\newcommand{\softneg}{\operatornamewithlimits{\tilde{\neg}}}
%--------------------------------------------------------

\begin{document}
\thispagestyle{empty}
\pagenumbering{roman}

\begin{center}
% TITLE
{\Large 
Automated Social Group Modeling for Characterizing and Forecasting Elections
}
\vfill
Aravindan Mahendiran
\vfill
Thesis submitted to the Faculty of the \\
Virginia Polytechnic Institute and State University \\
in partial fulfillment of the requirements for the degree of
\vfill
Master of Science \\
in \\
Computer Science and Applications
\vfill
Naren Ramakrishnan, Chair \\
Cal Ribbens \\
Aditya Prakash
\vfill
December 25, 2013 \\
Blacksburg, Virginia
\vfill
Keywords: Forecasting, Twitter, Query Expansion, Elections, Probabilistic Soft Logic \\
Copyright 2013, Aravindan Mahendiran
\end{center}
\pagebreak
\thispagestyle{empty}

\begin{center}
{\large Automated Social Group Modeling for Characterizing and Forecasting Elections}
\vfill
Aravindan Mahendiran
\vfill
(ABSTRACT)
\vfill
\end{center}
Twitter mining over the last few years has garnered a lot of attention from the research community. 
Strong correlations have been shown between Twitter \emph{‘mentions’} and stock markets, book sales and flu outbreaks which is then used for forecasting.
Even though such methodologies are accurate in forecasting the trends to a great extent, their performance is dictated by the domain specific vocabulary is used to track the relevant tweets. 
Such a vocabulary is usually provided by subject matter experts but is not exhaustive.
The language used in Twitter is drastically different from other forms of writing like those in news articles or even web blogs.  
It constantly evolves with time as users adopt popular hash-tags to express their opinion.
Thus, the vocabulary used by the forecasting algorithms needs to be dynamic in nature and should capture the rising trends of the domain. 
Otherwise, the prediction algorithms miss out on capturing the some of the most informative documents.
\newline 
We propose a novel unsupervised learning algorithm builds a vocabulary through modeling user preferences by exploiting the explicit and latent structure in such data sets.
We use Probabilistic Soft Logic, a framework for probabilistic reasoning over relational domains, to develop a query expansion algorithm that learns such a dynamic vocabulary for any given domain.  
Using 7 presidential elections from Latin America we show how such a query expansion methodology improves the recall and accuracy of two state of the art election prediction algorithms. 
Through this approach we achieve close to 2x increase in the number of documents used for predictions and 16.13\% reduction in the prediction error. 
\pagebreak

% Dedication and Acknowledgments are both optional
\chapter*{Dedication}
\vspace*{\fill}
\begin{center}
\Large\textit{To my awesomest brother, mother and friends}
\end{center}
\vspace{\fill}

\chapter*{Acknowledgments}
\vspace*{\fill}
\Large{First and foremost I would like to thank my advisor Dr.Naren Ramakrishnan. He was a mentor in every sense of the word and played a vital role in kindling my interest for data mining and machine learning.
\newline I would like to thank my colleagues Sathappan Muthiah, Rupinder Paul Khandpur. Without their support and help, I could not have completed my graduate study. 
}
\vspace{\fill}

\tableofcontents
\pagebreak

\listoffigures
\pagebreak

\listoftables
\pagebreak

\pagenumbering{arabic}
\pagestyle{myheadings}

%\doublespacing


\input{tex_files/Introduction.tex}
\input{tex_files/PSL.tex}
\input{tex_files/Prediction_Models.tex}
\input{tex_files/Results.tex}
\input{tex_files/Conclusions_FutureWork.tex}

%%%%%%%%%%%%%%%%%
%
% Include an EPS figure with this command:
%   \epsffile{filename.eps}
%

%%%%%%%%%%%%%%%%
%
% Do tables like this:

%  \begin{table}
%  \caption{The Graduate School wants captions above the tables.}
% \begin{center}
%  \begin{tabular}{ccc}
%  x & 1 & 2 \\ \hline
%  1 & 1 & 2 \\
%  2 & 2 & 4 \\ \hline
%  \end{tabular}
% \end{center}
%  \end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% If you are using BibTeX, uncomment the following:
% \thebibliography
%
% Otherwise, uncomment the following:
% \chapter*{Bibliography}

% \appendix

% In LaTeX, each appendix is a "chapter"
% \chapter{Program Source}

\cleardoublepage
%\phantomsection
\addcontentsline{toc}{chapter}{Bibliography}
\bibliographystyle{IEEEtran}
\markright{Aravindan Mahendiran \hfill References \hfill}
\bibliography{tex_files/references}
\end{document}
